### 2020-01-22 articletwo-sensitivityanalysis.Rmd
init

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

### Plot functions

<details>
  <summary>Open/Close</summary>

```{r Plot functions, code=readLines("articletwo-ggplot.R")}
source("articletwo-ggplot.R")
```

</details>

# Import previous session

```{r Import previous session}
pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
names.dset <- getdescriptions()
```

## Variables

```{r my variables}
var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
var.CL.min <- c("BL_AGE", "SEX")
var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
var.CL.opt <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB", "BP_TREAT")
permutations <- 99
```

# Sensitivity analysis

```{r sensitivity analysis bp treat}
adonis.species.bptreat <- calculate.betadiversity(pseq = pseq.species,
                                          matrix = bray.dist.m.species,
                                          vars = list("max" = var.CL.opt,
                                                      "min" = var.CL.min))

diversity.bptreat <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL.opt, "min" = var.CL.min),
                                 betadiversity = adonis.species.bptreat,
                                 names.dset = names.dset)

g.diversity.bptreat <- plot.diversities(diversity.bptreat)
ggsave(file = "cache/sensitivity-alphabeta-bptreat.png", plot = g.diversity.bptreat, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-bptreat.png"/>

```{r sensitivity analysis only healthy}
pseq.species.healthy <- subset_samples(pseq.species,
                                       PREVAL_DIAB == 0 &
                                       PREVAL_CHD == 0 &
                                       PREVAL_CR_ANYCANC == 0)

bray.dist.m.species.healthy <- bray.dist.m.species[pseq.species.healthy %>% meta %>% rownames,
                                                   pseq.species.healthy %>% meta %>% rownames]

adonis.species.healthy <- calculate.betadiversity(pseq = pseq.species.healthy,
                                          matrix = bray.dist.m.species.healthy,
                                          vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                      "min" = var.CL.min))

diversity.healthy <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                             "min" = var.CL.min),
                                 betadiversity = adonis.species.healthy,
                                 names.dset = names.dset)

g.diversity.healthy <- plot.diversities(diversity.healthy)
ggsave(file = "cache/sensitivity-alphabeta-healthy.png", plot = g.diversity.healthy, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-healthy.png"/>

## BMI

Normal weigth

```{r sensitivity analysis normal weight}
pseq.species.normalweight <- subset_samples(pseq.species, BMI > 18.5 & BMI <= 25)

bray.dist.m.species.normalweight <-
    bray.dist.m.species[pseq.species.normalweight %>% meta %>% rownames,
                        pseq.species.normalweight %>% meta %>% rownames]

adonis.species.normalweight <- calculate.betadiversity(pseq = pseq.species.normalweight,
                                          matrix = bray.dist.m.species.normalweight,
                                          vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                      "min" = var.CL.min))

diversity.normalweight <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                             "min" = var.CL.min),
                                 betadiversity = adonis.species.normalweight,
                                 names.dset = names.dset)

g.diversity.normalweight <- plot.diversities(diversity.normalweight)
ggsave(file = "cache/sensitivity-alphabeta-normalweight.png",
       plot = g.diversity.normalweight, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-normalweight.png"/>

```{r sensitivity analysis obese}
pseq.species.obese <- subset_samples(pseq.species, BMI > 30 & BMI <= 40)

bray.dist.m.species.obese <-
    bray.dist.m.species[pseq.species.obese %>% meta %>% rownames,
                        pseq.species.obese %>% meta %>% rownames]

adonis.species.obese <- calculate.betadiversity(pseq = pseq.species.obese,
                                          matrix = bray.dist.m.species.obese,
                                          vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                      "min" = var.CL.min))

diversity.obese <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                             "min" = var.CL.min),
                                 betadiversity = adonis.species.obese,
                                 names.dset = names.dset)

g.diversity.obese <- plot.diversities(diversity.obese)
ggsave(file = "cache/sensitivity-alphabeta-obese.png", plot = g.diversity.obese, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-obese.png"/>

```{r save session}
save.image(file = "session/session-sensitivity.Rdata")
```