### 2020-01-22 articletwo.Rmd

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)

dir.create("cache/", showWarnings = FALSE)
dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)

now <- format(Sys.time(), '%Y%m%d-%H%M%S')
```

```{r Command line arguments, include = FALSE, eval = FALSE}
file <- paste0("session/", sort(list.files("session"), decreasing = TRUE)[1])
message("Loading variables from file ", file)
load(file)
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(egg)
library(gridExtra)
library(png)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
library(DESeq2)
library(ggeffects)
library(pander)
library(hexbin)
```

Session info

```{r Session info}
pander(sessionInfo(), compact = TRUE)
```

# Sources

<details><summary>Functions</summary>

```{r Functions}
sourcefiles <- c("articletwo-officer.R",
                 "articletwo-rrbiome.R",
                 "articletwo-ggplot.R")
```

```{r import files, echo = FALSE}
for (f in sourcefiles) {
    source(f)
}
```

```{r embbed files, echo = FALSE}
xfun::embed_files(c("rrnmr.Rmd", sourcefiles))
```

</details>

# Loading data

Loading descriptions for clinical data

```{r variables, warning = FALSE}
names.dset <- getdescriptions()
```

```{r temp}
temp <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
temp %>% meta %>% nrow
```

Loading phyloseq object

```{r data}
pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.genus.coretaxa <- coretaxa(pseq.genus, detection = 0.1/100, prevalence = 1/100)
```

Species has average number of reads `r pseq.species %>% sample_sums %>% mean` and
genus `r pseq.genus %>% sample_sums %>% mean`. Core has length 
`r length(pseq.genus.coretaxa)`.

At species level meta has dimensions (`r dim(meta(pseq.species))`) and
there ntaxa is `r ntaxa(pseq.species)`. At genus level meta has
dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
`r ntaxa(pseq.genus)`.

# Variables

```{r my variables}
var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
var.CL.min <- c("BL_AGE", "SEX")
var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
```

# Heavy models

### Bray curtis distance matrix

```{r matrix calculation}
if (!file.exists("rds/bray.dist.m.species.rds")) {
    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
    saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
} else {
    bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
}
```

### Beta diversity

 ```{r adonis calculation}
 if (!file.exists("rds/adonis.species.rds")) {
     adonis.species <- calculate.betadiversity(pseq = pseq.species,
                                               matrix = bray.dist.m.species,
                                               vars = list("max" = var.CL,
                                                            "min" = var.CL.min))
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
 } else {
     adonis.species  <- readRDS("rds/adonis.species.rds")
 }
```

### PCoA

```{r pcoa calculate}
if (!file.exists("rds/pcoa.ordinate.rds")) {
    pcoa.abundances <- microbiome::transform(pseq.species, 'compositional')
    pcoa.ordinate <- ordinate(pcoa.abundances, method="PCoA", distance="bray")
    saveRDS(pcoa.ordinate, file = "rds/pcoa.ordinate.rds")
} else {
    pcoa.ordinate <- readRDS("rds/pcoa.ordinate.rds")
}
```

### DeSeq2

```{r deseq2}
if (!file.exists("rds/dds.rds") || !file.exists("rds/dds3.rds") || !file.exists("rds/ddsNa.rds")) {
    pseq.genus.core.deseq <-  prune_taxa(pseq.genus.coretaxa, pseq.genus)

    if (!file.exists("rds/dds.rds")) {
        dds <- lapply(c2l(var.BP), function(x, pseq) {
            dds.data <- phyloseq_to_deseq2(pseq, deseq.formula(x, var.CL))
            DESeq(dds.data,
                  test="Wald",
                  fitType="parametric",
                  parallel = TRUE,
                  BPPARAM=MulticoreParam(16))
        }, pseq = pseq.genus.core.deseq)
        saveRDS(dds, file = "rds/dds.rds")
    }
    if (!file.exists("rds/dds3.rds")) {
        dds3 <- lapply(deseq.list(var.CL, var.CL.min), function(x, pseq) {
            dds.data <- phyloseq_to_deseq2(pseq, deseq.formula(x, "HYPERTENSION"))
            DESeq(dds.data,
                  test="Wald",
                  fitType="parametric",
                  parallel = TRUE,
                  BPPARAM=MulticoreParam(16))
        }, pseq = pseq.genus.core.deseq)
        saveRDS(dds3, file = "rds/dds3.rds")
    }
    if (!file.exists("rds/ddsNa.rds")) {
        ddsNa <- mclapply(c2l(var.BP), function(x) {
            myDESeq(pseq = pseq.species,
                    vars = c(x, var.CL),
                    coreterm = "^Lactobacillus.*Bacteria")
        }, mc.cores = length(var.BP))
        saveRDS(ddsNa, file = "rds/ddsNa.rds")        
    }
} else {
    dds <- readRDS(file = "rds/dds.rds")
    dds3 <- readRDS(file = "rds/dds3.rds")
    ddsNa <- readRDS(file = "rds/ddsNa.rds")
}
```

# Results

## Characteristics

```{r Characteristics}
characteristicsTable(meta(pseq.species)) %>% kable
```

<details><summary>Salt subset</summary>

```{r Characteristics subset, echo = FALSE}
pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE) %>%
    meta %>%
    characteristicsTable %>%
    kable
```

</details>

## Microbial diversity

```{r alphabeta definitions}
diversity <- diversities(pseq = pseq.species,
                           vars = list("max" = var.CL, "min" = var.CL.min),
                           betadiversity = adonis.species,
                           names.dset = names.dset)
```

Results in minimum model of alpha diversity

```{r talbe alpha div min}
diversities.tidy(diversity) %>%
    kable
```

```{r save grob}
g.diversity <- plot.diversities(diversity)
ggsave(file = "cache/alpha-beta.png", plot = g.diversity, height = 6, width = 9)
```

<img src="cache/alpha-beta.png" />

## Principal coordinate analysis

```{r pcoa plot}
pcoa.plot <- pcoaplot(pcoa.ordinate)
ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.0, width=6, units = "in", dpi = 300)
```

<img src="cache/pcoa-species.png" />

## Associations for genera

```{r deseq continuous}
dset.deseq.signf <- deseqresults(dds, names.dset)
g.deseq <- deseqheatmap(dset.deseq.signf)
```

Number of significant associations `r sum(pull(dset.deseq.signf, qval) < 0.05)` for 
`r pull(dset.deseq.signf, Feature)  %>% unique %>% length` features.

### Number of associations per covariate

```{r results compare deseq2}
map(dds3, ~sum(results(.x, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)) %>%
    map_df(~as.data.frame(.x) %>% dplyr::rename(qval = .x), .id = "included") %>%
    kable
```

## Associations for Lactobacillus species

```{r lactobacillus species results}
dset.deseq.signf.na <- deseqresults(ddsNa, names.dset)
g.deseq.na <- deseqheatmap(dset.deseq.signf.na)
```

Number of significant associations `r sum(pull(dset.deseq.signf.na, qval) < 0.05)` for 
`r dset.deseq.signf.na %>% nrow` features. Number of
`r dset.deseq.signf.na %>% filter(log2FoldChange > 0) %>% nrow`
association were positive.

```{r deseq dual plot}
g.deseqgenus <- gtable_frame(ggplotGrob(g.deseq + theme(legend.position = "none")),
                  height = unit(length(unique((dset.deseq.signf$Feature))), 'mm'),
                  width = unit(25, 'mm'))

g.deseqspecies <- gtable_frame(ggplotGrob(g.deseq.na + theme(legend.position = c(0.7, 1.45))),
                  height = unit(length(unique((dds.results$Feature))), 'mm'),
                                width = unit(25, 'mm'))

g.deseq <- arrangeGrob(g.deseqgenus, g.deseqspecies, ncol = 2)

ggsave(file = "cache/deseq-full.png", plot = g.deseq, width = 14, height = 25, dpi = 300, unit = "cm")
```

<img src="cache/deseq-full.png" />

## Salt subset

## Association between genera and dU-NA

```{r lacto deseq calculation}
dds.salt <- myDESeq(pseq = pseq.genus,
                    coretaxa = pseq.genus.coretaxa,
                    vars = c("NA.", var.CL),
                    saltsubset = TRUE)
```

Lactobacillus prevalence in subsample

```{r lactobacillus prevalence}
pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
pseq_prevalence(pseq.genus.salt)
```

```{r significant results for salt subset}
dds.salt %>%
    results(., name = "NA.", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
    filter(row == "Lactobacillus (Bacteria)") %>%
    kable
```

Graphical representation for findings

```{r scatter plot from deseq}
g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
ggsave(file = "cache/gsalt.png", plot = g.salt, height = 3, width = 3, dpi = 300)
```

<img src="cache/gsalt.png" />

## Association between Lactobacilli and dU-NA

```{r species salt}
dds.salt.species <- myDESeq(pseq = pseq.species,
                        vars = c("NA.", var.CL),
                        coreterm = "^Lactobacillus.*Bacteria",
                        saltsubset = TRUE)
```

Significant results after FDR correction

```{r species salt results}
dds.salt.species %>%
    results(., name = "NA.", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
                  row = renametaxa(row)) %>%
    filter(qval < 0.05) %>%
    kable
```

# Supplements

Listing our "core" genera

```{r write supplement}
core.names.fourcols <- pseq.genus.coretaxa %>%
    renametaxa %>%
    sort %>%
    list.partition %>%
    mycbind

pseq.genus.coretaxa %>% length

core.names.flextable <- flextable(core.names.fourcols) %>%
    flextable::fontsize(size = 10, part = "header") %>%
    flextable::fontsize(size = 10, part = "body") %>%
    flextable::width(j=1:4, width = 1.8) %>%
    flextable::align(align = "left", part = "header") %>%
    flextable::align(align = "left")
core.names.flextable
```

Listing results for alpha diversity

```{r supplement table for alpha diversity}
alphadiversity.table <- diversity %>% map_df(~as.data.frame(.x), .id = "model") %>%
    mutate(mean_ci = sprintf("%.2f (%.2f - %.2f)", alpha.effect, alpha.low, alpha.high),
           pubp = pub.p(alpha.p)) %>%
    myspread(list = c2l("mean_ci", "pubp"), term = "Name", key = "model") %>%
    arrange(match(Name, c("Systolic BP", "Diastolic BP", "Mean arterial pressure", "Pulse pressure", "Hypertension"))) %>%
        select(Name, min_mean_ci, min_pubp, max_mean_ci, max_pubp)

typology.tbls2 <- data.frame(
    col_keys = colnames(alphadiversity.table),
    what = c("", rep("Age- and sex adjusted model", 2), rep("Multivaraible adjusted model", 2)),
    measure = c("", rep(c("beta (95%-CI)", "p"), 2)),
    stringsAsFactors = FALSE)

alphadiversity.flextable <-
    typologyformatter(data = alphadiversity.table, font = 12, typology = typology.tbls2) %>%
    flextable::width(j=1, width = 1.9) %>%
    flextable::width(j=c(2:2,4:4), width = 1.8) %>%
    flextable::width(j=c(3:3,5:5), width = 0.7)
alphadiversity.flextable 
```

```{r supplement table for beta diversity}
betadiversity.table <- diversity %>% map_df(~as.data.frame(.x), .id = "model") %>%
    mutate(rsquared = sprintf("%.3f%%", beta.R2*100),
           pubp = pub.p(beta.p)) %>%
    myspread(list = c2l("rsquared", "pubp"), term = "Name", key = "model") %>%
    arrange(match(Name, c("Systolic BP", "Diastolic BP", "Mean arterial pressure", "Pulse pressure", "Hypertension"))) %>%
    select(Name, min_rsquared, min_pubp, max_rsquared, max_pubp)

typology.tbls3 <- data.frame(
    col_keys = colnames(betadiversity.table),
    what = c("", rep("Age- and sex adjusted model", 2), rep("Multivaraible adjusted model", 2)),
    measure = c("", rep(c("R2", "p"), 2)),
    stringsAsFactors = FALSE)

betadiversity.flextable <-
    typologyformatter(data = betadiversity.table, font = 12, typology = typology.tbls3) %>%
    flextable::width(j=1, width = 1.9) %>%
    flextable::width(j=c(2:5), width = 0.8)

betadiversity.flextable
```

```{r supplement table deseq}
typology.tbls4 <- data.frame(
    col_keys = colnames(deseq.table),
    what = c("", rep("Systolic BP", 2), rep("Diastolic BP", 2),
             rep("Pulse pressure", 2), rep("Mean arterial pressure", 2), rep("Hypertension", 2)),
    measure = c("", rep(c("Log2FC±SE", "p"), 5)),
    stringsAsFactors = FALSE)

deseq.table <- dset.deseq.signf %>%
    mutate(lfc_se = sprintf("%.2f±%.2f", log2FoldChange, lfcSE),
           p.value = pub.p(qval)) %>%
    myspread %>%
    select(Feature, starts_with("Systolic"), starts_with("Diastolic"), starts_with("Pulse pressure"),
           starts_with("MAP"), starts_with("Hypertension"))

typologyformatter(data = deseq.table, font = 9, typology = typology.tbls4) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55)
#  %>%   save_as_html("cache/genus.html")
```

```{r supplement table deseq}
deseq.table.na <- dset.deseq.signf.na %>%
    mutate(lfc_se = sprintf("%.2f±%.2f", log2FoldChange, lfcSE),
           p.value = pub.p(qval)) %>%
    myspread %>%
    select(Feature, starts_with("Systolic"), starts_with("Diastolic"), starts_with("Pulse pressure"),
           starts_with("MAP"), starts_with("Hypertension"))

typologyformatter(data = deseq.table.na, font = 9, typology = typology.tbls4) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55) 
#  %>%   save_as_html("cache/species.html")
```

```{r save session}
save.image(file = paste0("session/session-", now, ".Rdata"))
```